---
title: "Getting Started with US Disease Tracker Data Standards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with US Disease Tracker Data Standards}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The US Disease Tracker Data Standards package provides comprehensive documentation,
templates, and validation tools for disease surveillance data. This vignette will
help you get started with using and contributing to the data standards.

## Package Structure

The package is organized into several key directories:

- **inst/standards/**: Formal data standard definitions
- **inst/templates/**: Templates for creating new standards
- **inst/examples/**: Example data files demonstrating compliance
- **data-raw/**: Data validation scripts

## Data Standards

### Case Reporting Standard

The case reporting standard defines the required and optional fields for disease
case reports. See `inst/standards/case-reporting-standard.md` for the complete
specification.

Required fields include:
- case_id
- report_date
- disease_code
- jurisdiction
- age
- sex

### Using Example Data

Example data files are provided to demonstrate proper data formatting:

```{r eval=FALSE}
# Path to example file
example_file <- system.file("examples", "case_report_example.json", 
                           package = "USDiseaseTrackerDocs")

# Read the example (requires jsonlite package)
# library(jsonlite)
# example_data <- fromJSON(example_file)
```

## Validating Data

Validation scripts help ensure data compliance with standards. Since validation
scripts are in `data-raw/`, they should be accessed from the source repository:

```{r eval=FALSE}
# If you have the repository cloned, source the validation script directly
# source("data-raw/validate_case_report.R")

# Validate your data
# result <- validate_case_report(your_data)
# 
# if (result$valid) {
#   message("Data is valid!")
# } else {
#   print(result$errors)
# }
```

## Creating New Standards

To create a new data standard:

1. Copy the template from `inst/templates/standard-definition-template.md`
2. Fill in all sections with your standard details
3. Create example data files in `inst/examples/`
4. Create a validation script in `data-raw/`
5. Submit a pull request for review

## Contributing

Contributions to the data standards are welcome! Please:

1. Review existing standards for consistency
2. Use the provided templates
3. Include examples and validation scripts
4. Follow the GitHub workflow for submissions

## Further Information

For more details, visit the [package website](https://usdiseasetracker.github.io/USDiseaseTracker-Docs/).
